---
params:
  outputs: NA
  lang: NA
output:
  html_document:
    css: styles.css
    includes:
      in_header: header.html
    anchor_sections: no
---

```{r setup, include = FALSE}
print_fr <- FALSE
print_en <- FALSE
print_pt <- FALSE
if (params$lang == "fr") {
  print_fr <- TRUE
  title <- "Naomi rapport de synthèse"
} else if (params$lang == "pt") {
  print_pt <- TRUE
  title <- "Naomi relatório de síntese"
} else {
  print_en <- TRUE
  title <- "Naomi comparison report"
}
```

---
title: `r title`
---

```{r read_outputs, echo=FALSE, message = FALSE, warning = FALSE}
#Set limit for scientific notation
options(scipen=10000000)

outputs <- params$outputs

# Read files if hintr rds provided
if(tolower(tools::file_ext(params$outputs)) %in% c("rds", "qs")) {

  model_object <- read_hintr_output(outputs)
  outputs <- model_object$output_package
  inputs <- readr::read_csv(model_object$info$inputs.csv, show_col_types = FALSE)
  options <- yaml::read_yaml(text = model_object$info$options.yml)
  packages <- readr::read_csv(model_object$info$packages.csv, show_col_types = FALSE)
  version <- packages[packages$name == "naomi", ]$version
  outputs$version <- version

  calibration_files <- names(model_object$info)
  if("calibration_options.yml" %in% calibration_files) {
    calibration_options <- yaml::read_yaml(text = model_object$info$calibration_options.yml)
  } else {
    # In the event there is no calibration options, add in defaults:
    calibration_options <- tibble::tribble( ~option, ~value,
                                    "spectrum_population_calibration", "none",
                                    "spectrum_plhiv_calibration_level","subnational",
                                    "spectrum_plhiv_calibration_strat","sex_age_group",
                                    "spectrum_artnum_calibration_level","national",
                                    "spectrum_artnum_calibration_strat","age_coarse",
                                    "spectrum_aware_calibration_level","national",
                                    "spectrum_aware_calibration_strat","age_coarse",
                                    "spectrum_infections_calibration_level","none",
                                    "spectrum_infections_calibration_strat","age_coarse",
                                    "calibrate_method","logistic")
  }
}

# Read files if zipped output package provided
if(grepl("\\.zip$", params$outputs)) {
  output_zip <- params$outputs
  outputs <- naomi::read_output_package(output_zip)
  inputs <- unz(output_zip, "info/inputs.csv")
  inputs <- read.csv(inputs)
  options <- unz(output_zip, "info/options.yml")
  options <- yaml::read_yaml(options)

  packages <- unz(output_zip, "info/packages.csv")
  packages <- readr::read_csv(packages, show_col_types = FALSE)
  version <- packages[packages$name == "naomi", ]$version
  outputs$version <- version



  if("calibration_options" %in% names(outputs$fit)) {
    calibration_options <- outputs$fit$calibration_options
  } else {
    calibration_options <- options[grep("calibration", names(options))]
  }
}

# Grab inputs and model options from model output zip file
# # concatenating strings where more than one option may be selected
report_date <- format(Sys.Date(), "%d-%m-%Y")
t1 <- options$calendar_quarter_t1
t2 <- options$calendar_quarter_t2
iso <- options$area_scope
quarter_t1 <- outputs$meta_period[outputs$meta_period$calendar_quarter == t1,]$quarter_label
quarter_t2 <- outputs$meta_period[outputs$meta_period$calendar_quarter == t2,]$quarter_label
level <- as.integer(options$area_level)
spectrum_file <- paste0(inputs[inputs$role == "pjnz",]$filename)


# Identify area_level_label for model estimates
area_level_map <- outputs$meta_area %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(area_level, area_level_label) %>%
  dplyr::summarise(.groups = "drop")

area_level_label <- area_level_map$area_level_label[area_level_map$area_level == level]

country <- paste0(outputs$meta_area$area_name[outputs$meta_area$area_id == iso],
                  sep = "", collapse = "")

# Determine if and when programme data has been included
include <- tibble::tibble(
  art = list(options$include_art_t1, options$include_art_t2),
  art_year = list(quarter_t1, quarter_t2),
  art_attend = list(options$artattend, options$artattend_t2)
) %>%
  dplyr::mutate_all(list(~dplyr::na_if(., "NULL"))) %>%
  dplyr::mutate_all(list(~dplyr::na_if(., "")))

# Select years where programme data is included
collapse_and <- " and "
if (print_fr) { collapse_and <- " et "} else if (print_pt) { collapse_and <- " e "}

# ART data included
if("true" %in% include$art) {
  df <- include %>% dplyr::filter(art == "true")
  art_year <- paste0(df$art_year, sep = "", collapse = collapse_and)
} else { art_year <- NA }

# ART attendance reallocation is implemented
if("true" %in% include$art_attend ) {
  df <- include %>% dplyr::filter(art_attend == "true")
  art_attend <- paste0(df$art_year, sep = "", collapse = collapse_and)
} else { art_attend <- NA }

# ANC ART programme data included
anc_art_options = c(options$anc_art_coverage_year1, options$anc_art_coverage_year2)
anc_art_options <- anc_art_options[!vlapply(anc_art_options, is_empty)]

if(length(anc_art_options)) {
  anc_art <- paste0(anc_art_options, sep = "", collapse = collapse_and)
} else { anc_art <- NA }

# ANC prev programme data
anc_prev_options = c(options$anc_prevalence_year1, options$anc_prevalence_year2)
anc_prev_options <- anc_art_options[!vlapply(anc_art_options, is_empty)]

if(length(anc_prev_options)) {
  anc_prev <- paste0(anc_prev_options, sep = "", collapse = collapse_and)
} else { anc_prev <- NA }


```


```{r, echo = FALSE, results = 'asis'}

if (print_en) {
  cat(paste0("## ", country, " Subnational HIV Estimates \n"))
  cat(paste0("Naomi estimates for **", quarter_t2, "** (report generated ", report_date, ")"))
}

if (print_fr) {
  cat(paste0("## ", country, " Subnational HIV Estimates \n"))
  cat(paste0("Naomi estimates for **", quarter_t2, "** (report generated ", report_date, ")"))
}

if (print_pt) {
  cat(paste0("## ", country, " Subnational HIV Estimates \n"))
  cat(paste0("Naomi estimates for **", quarter_t2, "** (report generated ", report_date, ")"))
}

```

::: {#translate lang="en"}

### Methods
[Naomi](https://onlinelibrary.wiley.com/doi/10.1002/jia2.25788) is a small-area estimation model for estimating HIV prevalence and PLHIV, ART coverage, and new HIV infections at district level by sex and five-year age group. The model combines district-level data about multiple outcomes from several sources in a Bayesian statistical model to produce robust indicators of subnational HIV burden.

:::



```{r, echo = FALSE, results = 'asis'}

if (print_en) {
  cat("### Comparison of Naomi model estimates to data inputs  \n")
  cat("\n### Household Survey \n")
}

if (print_fr) {
  cat("### Comparison of Naomi model estimates to data inputs  \n")
  cat("\n### Household Survey \n")
}

if (print_pt) {
  cat("### Comparison of Noami model estimates to data inputs  \n")
  cat("\n### Household Survey \n")
}

```

::: {#translate lang="en"}

Cross-sectional estimates for HIV prevalence, ART coverage, and HIV incidence are produced at the mid-point of the most recent nationally representative household survey. For HIV prevalence, the model is calibrated to survey data about HIV prevalence by subnational level, sex, and five-year age group from the most recent population-based survey (for example [Population HIV Impact Assessment survey](https://phia.icap.columbia.edu/) or [Demographic and Health Survey](https://dhsprogram.com/)).

:::

```{r, echo = FALSE, results = 'asis'}

if (print_en) {
  cat("This report compares Naomi estimates for ", quarter_t2, "to the following population-baes survey(s): \n")
  cat("\n### Household Survey \n")
}

if (print_fr) {
  cat("This report compares Naomi estimates for ", quarter_t2, "to the following population-baes survey(s): \n")
  cat("\n### Household Survey \n")
}

if (print_pt) {
  cat("This report compares Naomi estimates for ", quarter_t2, "to the following population-baes survey(s): \n")
  cat("\n### Household Survey \n")
}
```



```{r, echo = FALSE, results = 'asis'}
survey_prev <- paste0(options$survey_prevalence, collapse = ", ")
survey_art <- paste0(options$survey_art_coverage, collapse = ", ")
survey_recent <- paste0(options$survey_recently_infected, collapse = ", ")

text <- tibble::tibble(prefix = c(t_("PREVALENCE_SURVEY_PREFIX"),
                                  t_("ART_SURVEY_PREFIX")),
                        source = c(survey_prev,
                                   survey_art)) %>%
  dplyr::filter(source != "")
cat(paste0("* ", text$prefix, " _", text$source, "_"), sep = "\n")


```

::: {#translate lang="en"}
We anticipate that the Naomi estimates will be closely aligned with the survey estimates at a national level. Some variation is expected between model estimates and survey estimates at lower Admin levels where survey estimates may be sparse for certain age and sex groups.
:::

::: {#translate lang="fr"}
We anticipate that the Naomi estimates will be closely aligned with the survey estimates at a national level. Some variation is expected between model estimates and survey estimates at lower Admin levels where survey estimates may be sparse for certain age and sex groups.
:::

::: {#translate lang="pt"}
We anticipate that the Naomi estimates will be closely aligned with the survey estimates at a national level. Some variation is expected between model estimates and survey estimates at lower Admin levels where survey estimates may be sparse for certain age and sex groups.
:::

```{r, echo = FALSE, results = 'asis'}

if (print_en) {cat("\n### Geographic distribution \n")}
if (print_fr) {cat("\n### Geographic distribution \n")}
if (print_pt) {cat("\n### Geographic distribution \n")}

```

```{r, echo=FALSE, warning = FALSE, message = FALSE, out.width  =  "90%", results = 'asis'}

data <- outputs$inputs_outputs
calendar_quarter1 <- options$calendar_quarter_t1
calendar_quarter2 <- options$calendar_quarter_t2

#  Barplot prevalence
fig1 <- bar_plotly (data,
                    ind = "prevalence",
                    quarter = calendar_quarter1)

#  Barplot ART coverage
fig2 <- bar_plotly (data,
                    ind = "art_coverage",
                    quarter = calendar_quarter1)

htmltools::div(
  style = "display: flex; flex-direction: column",
  htmltools::div(fig1),
  htmltools::div(fig2),
)

```


```{r, echo=FALSE, warning = FALSE, message = FALSE, out.width  =  "90%", results = 'asis'}


survey_prev1 <- options$survey_prevalence[1]
survey_prev2 <- options$survey_prevalence[2]

#  Scatter plot survey prevalence
fig1 <-  scatter_plotly(data,
                       ind = "prevalence",
                       quarter = calendar_quarter1,
                       input_data = survey_prev1,
                       input_data_type = "survey")


#  Scatter plot survey ART coverage
fig2 <-  scatter_plotly(data,
                       ind = "art_coverage",
                       quarter = calendar_quarter1,
                       input_data = survey_art,
                       input_data_type = "survey")



if(is_empty(survey_prev2)) {

  # A single prevalence survey

  htmltools::div(
    style = "display: flex;",
    htmltools::div(fig1, style = "width: 50%;"),
    htmltools::div(fig2, style = "width: 50%"),
  )

} else {

  # Multiple prevalence surveys

  fig1B <-  scatter_plotly(data,
                       ind = "prevalence",
                       quarter = calendar_quarter1,
                       input_data = survey_prev2,
                       input_data_type = "survey")


  prev_plots <-   htmltools::div(
    style = "display: flex;",
    htmltools::div(fig1B, style = "width: 50%;"),
    htmltools::div(fig1, style = "width: 50%"),
  )

  htmltools::div(
    style = "display: flex; flex-direction: column",
    prev_plots,
    htmltools::div(fig2, style = "width: 50%")
  )
}

```

```{r, echo = FALSE, results = 'asis'}

if (print_en) {cat("\n### Age distribution \n")}
if (print_fr) {cat("\n### Age distribution \n")}
if (print_pt) {cat("\n### Age distribution \n")}

```

```{r, echo=FALSE, warning = FALSE, message = FALSE, out.width  =  "90%", results = 'asis'}


#  Barplot prevalence
fig1 <- age_bar_plotly (data,
                    ind = "prevalence",
                    quarter = calendar_quarter1)

#  Barplot ART coverage
fig2 <- age_bar_plotly (data,
                    ind = "art_coverage",
                    quarter = calendar_quarter1)

  htmltools::div(
    style = "display: flex; flex-direction: column",
    htmltools::div(fig1),
    htmltools::div(fig2)
    )

```
::: {#translate lang="en"}
*Static images of individual plots can be downloaded by hovering on the right hand corner*
:::

::: {#translate lang="fr"}
*Static images of individual plots can be downloaded by hovering on the right hand corner*
:::

::: {#translate lang="pt"}
*Static images of individual plots can be downloaded by hovering on the right hand corner*
:::

```{r, echo = FALSE, results= 'asis', include = print_en, eval = print_en}
programme_data <- c(art_year, anc_prev, anc_art)

if (sum(is.na(programme_data)) <3 ) {

   cat(paste0("\n### Routinely collected programme data: \n", sep = "\n"))

   cat(paste0("For the HIV prevalence and ART coverage components of the model, ART service delivery numbers inform subnational estimates for the number of PLHIV. Since the Household survey sample size in each district is relatively small, routinely reported data about HIV prevalence among pregnant women attending their first antenatal care visit, extracted from the national health information system, are used to improve estimates of the spatial pattern of HIV. \n" , sep = "\n"))


   cat(paste0("\nThis report compares Naomi estimates to: \n", sep = "\n"))

  text2 <- tibble::tibble(prefix = c("National programme data on numbers on ART for ",
                                  "National programme data on ANC HIV prevalence for ",
                                  "National programme data on ANC ART coverage for "
                                  ),
                       source = c(art_year,
                                  anc_prev,
                                  anc_art)) %>% tidyr::drop_na()

  cat(paste0("\n ",cat(paste0("* ", text2$prefix, "_", text2$source, "_"), sep = "\n")))

 cat(paste0("The plots below compare ANC programme to age and sex matched estimates of the general population produced by Naomi. While these data should be positively correlated, we expect higher HIV prevalence and ART coverage in all females ages 15-49 relative to women attending antenatal services."  , sep = "\n"))

}


```

```{r, echo = FALSE, results= 'asis', include = print_fr, eval = print_fr}
programme_data <- c(art_year, anc_prev, anc_art)

if (sum(is.na(programme_data)) <3 ) {

  cat(paste0("\n### Données du programme opérationnel: \n", sep = "\n"))

  text2 <- tibble::tibble(prefix = c("Données national sur la couverture TARV pour ",
                                  "Données national sur la prévalence du VIH parmi CPN pour ",
                                  "Données national sur la couverture du TARV parmi CPN pour "
                                  ),
                       source = c(art_year,
                                  anc_prev,
                                  anc_art)) %>% tidyr::drop_na()

  cat(paste0("\n ",cat(paste0("* ", text2$prefix, "_", text2$source, "_"), sep = "\n")))

}

```


```{r, echo = FALSE, results= 'asis', include = print_pt, eval = print_pt}
programme_data <- c(art_year, anc_prev, anc_art)

if (sum(is.na(programme_data)) <3 ) {

  cat(paste0("\n### Dados do programa operacional: \n", sep = "\n"))

  text2 <- tibble::tibble(prefix = c("Dados do programa nacional sobre a cobertura de TARV para ",
                                  "Dados do programa nacional sobre a prevalência de VIH na CPN para ",
                                  "Dados do programa nacional sobre a cobertura de TARV na CPN para "
                                  ),
                       source = c(art_year,
                                  anc_prev,
                                  anc_art)) %>% tidyr::drop_na()

  cat(paste0("\n ",cat(paste0("* ", text2$prefix, "_", text2$source, "_"), sep = "\n")))



}

```


```{r, echo=FALSE, warning = FALSE, message = FALSE, out.width  =  "90%", results = 'asis'}

#-------------------------------------------------------------------------------
# ANC data: interactive scatter plot
#-------------------------------------------------------------------------------
anc_t1 <- outputs$fit$data_options$anc_prev_year_t1
anc_t2 <- outputs$fit$data_options$anc_prev_year_t2
has_anc <- !is_empty(anc_t1) && !is_empty(anc_t2)

if(has_anc) {
  #  Scatter plot ANC prevalence
  fig1 <-  scatter_plotly(data,
                          ind = "anc_prevalence_age_matched",
                          quarter = calendar_quarter1,
                          input_data = paste0("ANC ", anc_t1) ,
                          input_data_type = "ANC programme",
                          sex_disag = "female")

  fig2 <-  scatter_plotly(data,
                          ind = "anc_prevalence_age_matched",
                          quarter = calendar_quarter2,
                          input_data = paste0("ANC ", anc_t2) ,
                          input_data_type = "ANC programme",
                          sex_disag = "female")


  # Scatter plot ANC ART coverage

  fig3 <-  scatter_plotly(data,
                          ind = "anc_art_coverage_age_matched",
                          quarter = calendar_quarter1,
                          input_data = paste0("ANC ", anc_t2),
                          input_data_type = "ANC programme",
                          sex_disag = "female")

  fig4 <-  scatter_plotly(data,
                          ind = "anc_art_coverage_age_matched",
                          quarter = calendar_quarter2,
                          input_data = paste0("ANC ", anc_t2),
                          input_data_type = "ANC programme",
                          sex_disag = "female")


  prev_plots <-   htmltools::div(
    style = "display: flex;",
    htmltools::div(fig1, style = "width: 50%;"),
    htmltools::div(fig2, style = "width: 50%;")
  )

  art_plots <-   htmltools::div(
    style = "display: flex;",
    htmltools::div(fig3, style = "width: 50%;"),
    htmltools::div(fig4, style = "width: 50%;")
  )

  htmltools::div(
    style = "display: flex; flex-direction: column",
    prev_plots,
    art_plots
  )
}
```




::: {#translate lang="en"}
_Version_

The Naomi model is supported by [UNAIDS](https://www.unaids.org/en) and developed and maintained by the [MRC Centre for Global Infectious Disease Analysis](https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis) at Imperial College London. The model receives technical guidance from the [UNAIDS Reference Group on Estimates, Modelling, and Projections](http://epidem.org/). The model was first used in 2020 and continues to be developed responsive to new data and HIV strategic information needs.
:::
