---
title: "Naomi Model Workflow Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Naomi Model Workflow Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
                    collapse = TRUE,
                    comment = "#>"
                  )
unlink("outputs", recursive = TRUE)
```




```{r setup, message = FALSE}
library(naomi)
library(tidyverse)
library(sf)
```

# 0. Prepare webtool GeoJSON input

The MVP version of Naomi web tool allows upload of a single GeoJSON file for
specifying the area hierarchy. This preprocessing step joins the area tables
into a single long format dataset and saves as a GeoJSON for upload to the
web tool.

```{r load_area_data, message = FALSE}

area_levels <- read_csv(system.file("extdata/areas/area_levels.csv", package = "naomi"))
area_hierarchy  <- read_csv(system.file("extdata/areas/area_hierarchy.csv", package = "naomi"))
area_boundaries <- sf::read_sf(system.file("extdata/areas/area_boundaries.geojson", package = "naomi"))

area_merged <- area_hierarchy %>%
  left_join(
    area_levels %>% select(area_level, area_level_label, display, naomi_level)
  ) %>%
  left_join(
    area_boundaries
  )

```
```{r message = FALSE}
st_write(area_merged, file.path(tempdir(), "area_merged.geojson"), delete_dsn = TRUE)
```

# 1. (Up)Load data inputs

Area hierarchy and boundaries

```{r }
area_merged <- read_sf(file.path(tempdir(), "area_merged.geojson"))

areas <- create_areas(area_merged = area_merged)
```

Population data

```{r load_population_data, message = FALSE}
pop_agesex <- read_csv(system.file("extdata/population/population_agesex.csv", package = "naomi"))
```

Survey data

```{r load_survey_data, message = FALSE}
survey_hiv_indicators <- read_csv(system.file("extdata/survey/survey_hiv_indicators.csv", package = "naomi"))
```

Programme data


```{r message = FALSE}
art_number <- read_csv(system.file("extdata/programme/art_number.csv", package = "naomi"))
anc_testing <- read_csv(system.file("extdata/programme/anc_testing.csv", package = "naomi"))
```

Use only adult ART data for now

```{r }
art_number <- art_number %>%
  filter(age_group == "15+")
```

Programme data

Spectrum PJNZ

```{r }
pjnz <- system.file("extdata/mwi2019.PJNZ", package = "naomi")
spec <- extract_pjnz_naomi(pjnz)
```

# 2. Choose model areas and time points

The following are required to be provided to define the model state space:

* `scope`: A collection of `area_id`s defining the set of areas to be modelled.
   Usually this is simply national level, so the level 0 `area_id`.
* `level`: Area level at which to fit model.
* `quarter_id_t1`: The first time point for the model--approximately the midpoint
  of the household survey data used.
* `quarter_id_t2`: The second time point for the model--the current time for which
   estimates are needed.

```{r }
scope <- "MWI"
level <- 4
calendar_quarter_t1 <- "CY2016Q1"
calendar_quarter_t2 <- "CY2018Q3"
```

The following select data inputs to model fitting from the uploaded datasets.
Providing `NULL` for any will exclude that data source from model fitting.

* Multiple household survey may be used in fitting, but they must be rougly
  contemporaneous around `quarter_id_t1`.
* Only survey ART coverage or survey VLS should be included from a given survey,
  not both. ART coverage is preferred if both are available.
* `artnum_quarter_id_t1` and `artnum_quarter_id_t1` are the time point at
  which current on ART programme data will be used to estimte ART coverage.
  They are typically the same `quarter_id_t1` and `quarter_id_t2` if ART
  programme data are used.
* `anc_quarter_id_t1` and `anc_quarter_id_t2` are typically a range of 3-4 quarters.    Data will be aggregated over these quarters for a larger sample size. They
  will typically be consecutive quarters, though a quarter could be dropped for
  example if there were reporting problems known to affect a given quarter.
Survey IDs to include in fitting

```{r }
prev_survey_ids  <- c("MWI2016PHIA", "MWI2015DHS")
artcov_survey_ids  <- "MWI2016PHIA"
vls_survey_ids <- NULL
recent_survey_ids <- "MWI2016PHIA"

artnum_calendar_quarter_t1 <- "CY2016Q1"
artnum_calendar_quarter_t2 <- "CY2018Q3"

anc_prevalence_year1 <- 2016
anc_prevalence_year2 <- 2018

anc_art_coverage_year1 <- 2016
anc_art_coverage_year2 <- 2018
```

# 3. Review input data

# 4. Prepare model inputs
Setup the model

```{r }
naomi_mf <- naomi_model_frame(areas,
                              pop_agesex,
                              spec,
                              scope = scope,
                              level = level,
                              calendar_quarter_t1,
                              calendar_quarter_t2)
```

Prepare data inputs

```{r }
naomi_data <- select_naomi_data(naomi_mf,
                                survey_hiv_indicators,
                                anc_testing,
                                art_number,
                                prev_survey_ids,
                                artcov_survey_ids,
                                recent_survey_ids,
                                vls_survey_ids,
                                artnum_calendar_quarter_t1,
                                artnum_calendar_quarter_t2,
                                anc_prevalence_year1,
                                anc_prevalence_year2,
                                anc_art_coverage_year1,
                                anc_art_coverage_year2)
```

5. Fit model
Prepare model inputs and initial parameters

```{r }
tmb_inputs <- prepare_tmb_inputs(naomi_data)
```

Fit the TMB model

```{r fit_model, cache = TRUE}
fit <- fit_tmb(tmb_inputs)
```

Calculate model outputs. We can calculate outputs based on posterior mode
estimates before running `report_tmb()` to calculate posterior intervals.

```{r }
outputs <- output_package(fit, naomi_mf, areas)
```

The output package consists of a data frame of indicators and metadata
defining the labels for each indicator.

```{r }
names(outputs)
```

If uncertainty has not been calcualted yet, the output object retures values
for `mode`, but not `mean` or `lower` and `upper` 95% uncertainty ranges.

```{r }
outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",
    age_group == "15-49"   # Age group 15-49
  ) %>%
  head()
```

The function `add_output_labels()` returns the indicators table
with labels added as additional columns.

```{r }
add_output_labels(outputs) %>%
  dplyr::filter(
    indicator == "prevalence",  # HIV prevalence
    age_group == "15-49"   # Age group 15-49
  ) %>%
  head()
```

Calculate uncertainty ranges and add to the output object
(This is time consuming and memory intensive.

```{r sample_outputs, cache = TRUE}
system.time(fit <- sample_tmb(fit))
```

Regenerate outputs with uncertainty ranges.

```{r make_output_package, cache = TRUE}
system.time(outputs <- output_package(fit, naomi_mf, areas))

outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",
    age_group == "15-49"
  ) %>%
  head()
```

Save model outputs to ZIP

```{r save_outputs, message = FALSE, results = "hide"}
dir.create("outputs", showWarnings = FALSE)
save_output_package(outputs, "mwi_outputs", "outputs", with_labels = FALSE)
save_output_package(outputs, "mwi_outputs_with_labels", "outputs", with_labels = TRUE)
save_output_package(outputs, "mwi_outputs_single_csv", "outputs", with_labels = TRUE, single_csv = TRUE)
save_output_package(outputs, "mwi_outputs_single_csv_unlabelled", "outputs", with_labels = FALSE, single_csv = TRUE)


## #' 6. Plot some model outputs

indicators <- add_output_labels(outputs) %>%
  left_join(outputs$meta_area %>% select(area_level, area_id, center_x, center_y)) %>%
  sf::st_as_sf()
```

15-49 prevalence by district

```{r prev_by_district_15, fig.height = 4, fig.width = 7}
indicators %>%
  filter(age_group == "15-49",
         indicator == "prevalence",
         area_level == 4) %>%
  ggplot(aes(fill = mode)) +
  geom_sf() +
  viridis::scale_fill_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

15-49 prevalence by Zone


```{r prev_by_zone_15, fig.height = 4, fig.width = 7}
indicators %>%
  filter(age_group == "15-49",
         ## sex == "both",
         indicator == "prevalence",
         area_level == 2) %>%
  ## semi_join(get_area_collection(areas, level = 3, area_scope = "MWI.3")) %>%
  ggplot(aes(fill = mean)) +
  geom_sf() +
  viridis::scale_fill_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

Age-specific prevalence, national

```{r age_specific_prev, fig.height = 5, fig.width = 7}
indicators %>%
  left_join(outputs$meta_age_group) %>%
  dplyr::filter(area_level == 0,
         sex != "both",
         age_group_id %in% 1:17,
         indicator == "prevalence") %>%
  left_join(get_age_groups()) %>%
  mutate(age_group = fct_reorder(age_group_label, age_group_id)) %>%
  ggplot(aes(age_group, mean, ymin = lower, ymax = upper, fill = sex)) +
  geom_col(position = "dodge") +
  geom_linerange(position = position_dodge(0.8)) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent_format(1)) +
  facet_wrap(~area_name) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5))
```

15-64 ART coverage by district

```{r art_cov_district, fig.height = 4, fig.width = 7}
indicators %>%
  filter(age_group_id == "15-64",
         area_level == 4,
         indicator_id == "art_coverage") %>%
  ggplot(aes(fill = mean)) +
  geom_sf() +
  viridis::scale_fill_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

Age-specific ART coverage, national

```{r age_specific_art_cov, fig.height = 5, fig.width = 7}
indicators %>%
  left_join(outputs$meta_age_group) %>%
  dplyr::filter(area_level == 0,
         sex != "both",
         age_group_id %in% 1:17,
         indicator == "art_coverage") %>%
  left_join(get_age_groups()) %>%
  mutate(age_group = fct_reorder(age_group_label, age_group_id)) %>%
  ggplot(aes(age_group, mean, ymin = lower, ymax = upper, fill = sex)) +
  geom_col(position = "dodge") +
  geom_linerange(position = position_dodge(0.8)) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent_format(1)) +
  facet_wrap(~area_name) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5))
```

ART coverage by age/sex and region


```{r art_cov_age_sex, fig.height = 4, fig.width = 7}
indicators %>%
  left_join(outputs$meta_age_group) %>%
  filter(area_level == 1,
         sex != "both",
         age_group_id %in% 1:17,
         indicator == "art_coverage") %>%
  left_join(get_age_groups()) %>%
  mutate(age_group = fct_reorder(age_group_label, age_group_id)) %>%
  ggplot(aes(age_group, mean, ymin = lower, ymax = upper, fill = sex)) +
  geom_col(position = "dodge") +
  geom_linerange(position = position_dodge(0.8)) +
  facet_wrap(~area_name) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5))
```

Bubble plot prevalence and PLHIV


```{r bubble_plot, fig.height = 4, fig.width = 7}
indicators %>%
  filter(age_group == "15-64",
         area_level == 4,
         indicator %in% c("prevalence", "plhiv")) %>%
  select(sex, center_x, center_y, indicator_label, mean) %>%
  spread(indicator_label, mean) %>%
  ggplot() +
  geom_sf() +
  geom_point(aes(center_x, center_y, colour = `HIV Prevalence`, size = PLHIV)) +
  viridis::scale_color_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

